---
title: "homework-03"
output: html_document
---

https://github.com/Zach-Tech/ENVS-193DS_homework-03.git


```{r setup}
# Load required libraries
library(tidyverse)
library(here)
library(gt)
library(janitor)
library(readxl)
```

## Problem 1a. Data Summarizing (5 points)

To compare my response variable—**number of steps per day**—between different **days of the week**, I could calculate the **mean number of steps** for each day. This would allow me to identify patterns, such as whether I am more active on weekdays compared to weekends.

This comparison is informative because my schedule varies during the week. For example, I have back-to-back classes on Monday and Wednesday, which usually requires more walking, whereas I tend to stay home on Sundays, resulting in fewer steps.

# Load the data
```{r}
data <- read_csv(here("data", "personal_data.csv"))
```

## Problem 1b. Visualization (10 points)

```{r}
# Summarize: calculate mean steps by weekend vs weekday
library(dplyr)
library(ggplot2)

steps_summary <- data %>%
  group_by(is_weekend) %>%
  summarize(mean_steps = mean(steps), .groups = "drop")

# Plot: points for individual observations + bar for mean
ggplot(data, aes(x = is_weekend, y = steps)) +
  geom_jitter(width = 0.15, size = 3, alpha = 0.7, color = "#1f78b4") +  # individual data
  geom_col(data = steps_summary, aes(x = is_weekend, y = mean_steps), 
           fill = "#33a02c", alpha = 0.5, width = 0.6) +  # mean bars
  labs(
    title = "Comparison of Daily Step Count: Weekdays vs. Weekends",
    x = "Day Type",
    y = "Number of Steps"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold")
  )
```
